Modeling the Food Insecurity Experience Scale (FIES) in the Present and Future

### 2020-09-02 docs/modeling_fies.tex
Notes of Covariates

\author[1]{Matthew Cooper}
\author[2]{Benjamin Müller}

\affil[1]{T.H. Chan School of Public Health, Harvard University}
\affil[2]{TU Wien}

@@

To disaggregate the FIES by urban-rural areas, we used data on urbanization from \cite{Jiang2017}.

@@

\begin{table}[H]
	\begin{tabular}{llll}
		\toprule
		Name & Availability & Scale & Source \\
		\midrule
		Agriculture As A Percentage Of GDP & Annual & National & \cite{TheWorldBank2016} \\
		Net Oda Per Capita & Annual & National & \cite{TheWorldBank2016} \\
		Percent of Area Bare Ground & Annual & Admin-1 & \cite{Song2018} \\
		Percent Of Area With Built Up Land Cover & 2000 \& 2014 & Admin-1 & \cite{Pesaresi2015} \\
		Staple Crop Proction Per Capita & Annual & National & \cite{FAOSTAT2018} \\
		Elevation & Static & Admin-1 & \cite{USGS1996} \\
		Net School Enrollment & Annual & National & \cite{TheWorldBank2016} \\
		Forest Cover & Annual & Admin-1 & \cite{Song2018} \\
		Government Effectiveness Indicator & Annual & National & \cite{Kaufmann2011} \\
		GDP PPP Per Capita & Annual & Admin-1 & \cite{Kummu2018} \\
		HDI & Annual & Admin-1 & \cite{Kummu2018} \\
		Value Of All Imports Per Capita & Annual & National & \cite{TheWorldBank2016} \\
		Irrigated Area & 2000 & Admin-1 & \cite{SiebertS.DollP.FeickS.FrenkenK.Hoogeveen2013} \\
		Time To Travel To A City  & 2015 & Admin-1 & \cite{Uchida2008} \\
		Mean Annual Precipitation & Static & Admin-1 & \cite{Funk2015} \\
		NDVIi & Annual & Admin-1 & \cite{Song2018} \\
		People Per Pixel & 2010, 2015, 2020 & Admin-1 & \cite{Doxsey-Whitfield2015} \\
		Topographic Roughness Index & Static & Admin-1 & \cite{USGS1996, Riley1999} \\
		Political Stability Index & Annual & Admin-1 & \cite{Kaufmann2011} \\
		Number Of Conflict Deaths Within 50Km & Annual & Admin-1 & \cite{Eriksson2015} \\
		Nutritional Diversity Of Agriculture & 2015 & Admin-1 & \cite{Herrero2017a} \\
		Average Maximum Temperature & Roughly Static & Admin-1 & \cite{Sheffield2006} \\
		Number Of Buffaloes Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Cattle Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Chickens Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Ducks Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Horses Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Pigs Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Sheep Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Incidence \textit{P. falciparum} Malaria & Annual & Admin-1 & \cite{Weiss2019} \\
		Incidence \textit{P. vivax} Malaria & Annual & Admin-1 & \cite{Weiss2019} \\
		\bottomrule
	\end{tabular}
	\caption{Co-Variates Included in Present-Day Analysis}
	\label{tab:covars1}
\end{table}

\subsection{Covariates Available in the Present and Future}
To model the FIES in the future, we use projections derived in line with the Shared Socio-Economic Pathways (SSPs).  The SSPs are a framework to estimate what the future might look like in terms of human development and climate-change mitigation, based on a number of scenarios.  We use projections of GDP \cite{Dellink2017}, education \cite{KC2017}, and urbanization \cite{Jiang2017} for SSP2, the middle-of the road pathway.

>>>>

\author[1]{Matthew Cooper}
\author[2]{Benjamin Müller}
\author[3]{Juan Carlos Laso}
\author[4]{Homi Kharas}

\affil[1]{T.H. Chan School of Public Health, Harvard University}
\affil[2]{Department of Economics, Vienna University of Economics and Business}

@@

To disaggregate the FIES by urban-rural areas, we used data on urbanization from \cite{Jiang2017}.
%once implemented add here more to disaggregation

@@

%NOTE: WHEN JUSTIFYING COVARIATES, SAY THAT WE USE MANY FROM DEVELOPING COUNTRIES BECAUSE THE CAUSES OF FOOD INSECURITY ARE MORE DIVERSE THERE: ALMOST A KARENINA PRINCIPLE

%update/create this table once we have our final model
\begin{table}[H]
	\begin{tabular}{llll}
		\toprule
		Name & Availability & Scale & Source \\
		\midrule
		Agriculture As A Percentage Of GDP & Annual & National & \cite{TheWorldBank2016} \\
		Net Oda Per Capita & Annual & National & \cite{TheWorldBank2016} \\
		Percent of Area Bare Ground & Annual & Admin-1 & \cite{Song2018} \\
		Percent Of Area With Built Up Land Cover & 2000 \& 2014 & Admin-1 & \cite{Pesaresi2015} \\
		Staple Crop Proction Per Capita & Annual & National & \cite{FAOSTAT2018} \\
		Elevation & Static & Admin-1 & \cite{USGS1996} \\
		Net School Enrollment & Annual & National & \cite{TheWorldBank2016} \\
		Forest Cover & Annual & Admin-1 & \cite{Song2018} \\
		Government Effectiveness Indicator & Annual & National & \cite{Kaufmann2011} \\
		GDP PPP Per Capita & Annual & Admin-1 & \cite{Kummu2018} \\
		HDI & Annual & Admin-1 & \cite{Kummu2018} \\
		Value Of All Imports Per Capita & Annual & National & \cite{TheWorldBank2016} \\
		Irrigated Area & 2000 & Admin-1 & \cite{SiebertS.DollP.FeickS.FrenkenK.Hoogeveen2013} \\
		Time To Travel To A City  & 2015 & Admin-1 & \cite{Uchida2008} \\
		Mean Annual Precipitation & Static & Admin-1 & \cite{Funk2015} \\
		NDVIi & Annual & Admin-1 & \cite{Song2018} \\
		People Per Pixel & 2010, 2015, 2020 & Admin-1 & \cite{Doxsey-Whitfield2015} \\
		Topographic Roughness Index & Static & Admin-1 & \cite{USGS1996, Riley1999} \\
		Political Stability Index & Annual & Admin-1 & \cite{Kaufmann2011} \\
		Number Of Conflict Deaths Within 50Km & Annual & Admin-1 & \cite{Eriksson2015} \\
		Nutritional Diversity Of Agriculture & 2015 & Admin-1 & \cite{Herrero2017a} \\
		Average Maximum Temperature & Roughly Static & Admin-1 & \cite{Sheffield2006} \\
		Number Of Buffaloes Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Cattle Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Chickens Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Ducks Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Horses Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Pigs Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Number Of Sheep Per Square Km & 2010 & Admin-1 & \cite{Robinson2014} \\
		Incidence \textit{P. falciparum} Malaria & Annual & Admin-1 & \cite{Weiss2019} \\
		Incidence \textit{P. vivax} Malaria & Annual & Admin-1 & \cite{Weiss2019} \\
		\bottomrule
	\end{tabular}
	\caption{Co-Variates Included in Present-Day Analysis}
	\label{tab:covars1}
\end{table}

\begin{table}[H]
	\begin{tabular}{llll}
		\toprule
		Name & Availability & Scale & Source \\
		\midrule
        HDI & Annual & Subnational & GDL \\
        Life Expectancy & Annual & Subnational & GDL \\
        GNI Per Capita & Annual & Subnational & GDL \\
        Expected Years School & Annual & Subnational & GDL \\
        Mean Years School & Annual & Subnational & GDL \\
        Country Programmable Assistance (CPA) & Annual & National & OECD Database \\
        AG PCT GD & Annual & National & World Bank \\
        Import Value Per Capita & Annual & National & World Bank \\
        Undernourishment & Annual & National & World Bank, FAO \\
        government_effectiveness & Annual & National & World Bank WGIs \\
        control_corruption & Annual & National & World Bank WGIs \\
        regulatory_quality & Annual & National & World Bank WGIs \\
        rule_of_law & Annual & National & World Bank WGIs \\
        stability_noviolence & Annual & National & World Bank WGIs \\
        voice_accountability & Annual & National & World Bank WGIs \\
        mean annual precip & static & subnational & WorldClim \\
        mean average temp & static & subnational & WorldClim \\
        forest cover & 2016 & subnational & Song et al \\
        elevation & static & subnational & ETOPO5 \\
        ruggedness & static & subnational & ETOPO5 \\

        %TODO:
        %Nutritional Diversity
        %Malaria
        %Stunting
        %Livestock
        %Buitup
        %Irrigation
        %TTC
        %Sum of conflict deaths
        %Urban-Rural pop

		\bottomrule
	\end{tabular}
	\caption{Co-Variates Included in Present-Day Analysis NEW}
	\label{tab:covars2}
\end{table}

\subsection{Covariates Available in the Present and Future}
To model the FIES in the future, we use projections derived in line with the Shared Socio-Economic Pathways (SSPs).  The SSPs are a framework to estimate what the future might look like in terms of human development and climate-change mitigation, based on a number of scenarios.  We use projections of GDP \cite{Dellink2017}, education \cite{KC2017}, and urbanization \cite{Jiang2017} for SSP2, the middle-of the road pathway.
%in general we use middle of the road but maybe also other scenarios to account for uncertainty
